---
title: "Turtlel deaths estimation using Richards functions"
output: word_document
---

This document describes the method of estimating the total deaths of green turtles at Raine Islandj using counts from a few days over each nesting season. This is the same approach as when the total hatchling production was estimated. 

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(lubridate)
library(readr)
library(reshape2)
library(jagsUI)
library(bayesplot)

#source("Dunstan_functions.R")

col.def <- cols(Date = col_date(format = "%m/%d/%Y"),
                Heat = col_integer(),
                Cliff = col_integer(),
                Total = col_integer())

# Begin and end dates for each season
Day.begin <- "10-15"
Day.end <- "04-30"
# define nesting seasons - not the same as calendar years
# also define day of season, starts on 10-15

data.1 <- read_csv(file = "data/Turtle_deaths_with_zeros.csv",
                   col_types = col.def) %>% 
  mutate(Year = year(Date),
         DOY = yday(Date),
         Season = ifelse(Date >= paste0(Year, "-", Day.end), 
                         Year, Year-1),
         Season.f = as.factor(Season), 
         DOS = as.numeric(Date - as.Date(paste0(Season, "-", Day.begin)))) %>%  
  group_by(Season.f) %>%
  mutate(Cum_Heat = cumsum(Heat),
         Cum_Cliff = cumsum(Cliff))

```

Take a look at data first. Time series of heat-caused deaths:

```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = Date, y = Heat),
             color = "red") + 
  geom_path(aes(x = Date, y = Heat),
            color = "red") + 
    geom_point(aes(x = Date, y = Cliff),
               color = "blue") + 
  geom_path(aes(x = Date, y = Cliff),
            color = "blue", alpha = 0.6) +
  labs(title = "Heat and Cliff fall")


# n.days <- as.numeric(as.Date(paste0("2018-", Day2)) - as.Date(paste0("2017-", Day1)))
# 
# seasons <- unique(data.1$Season)
# 
# data.1 %>% mutate(DOSeason = as.numeric(Date - as.Date(paste0(Season, "-", Day1)))) -> data.3
```

```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = Date, y = log(Heat)),
             color = "red") + 
  geom_path(aes(x = Date, y = log(Heat)),
            color = "red") + 
    geom_point(aes(x = Date, y = log(Cliff)),
               color = "blue") + 
  geom_path(aes(x = Date, y = log(Cliff)),
            color = "blue", alpha = 0.6) + 
  labs(title = "Heat and cliff fall")
```


```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, 
                 y = (Heat))) + 
  geom_path(aes(x = DOS,
                y = (Heat))) +
  facet_wrap(. ~ Season)

```



```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, 
                 y = (Cliff))) + 
  geom_path(aes(x = DOS,
                y = (Cliff))) +
  facet_wrap(. ~ Season)

```



```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, 
                 y = log(Heat+1))) + 
  geom_path(aes(x = DOS,
                y = log(Heat+1))) +
  facet_wrap(. ~ Season)

```



```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, 
                 y = log(Cliff+1))) + 
  geom_path(aes(x = DOS,
                y = log(Cliff+1))) +
  facet_wrap(. ~ Season)

```


```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, y = Heat, color = Season.f)) + 
  geom_path(aes(x = DOS, y = Heat, color = Season.f)) +
  theme(legend.position = "none")
```


```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, y = Cliff, color = Season.f)) + 
  geom_path(aes(x = DOS, y = Cliff, color = Season.f)) +
  theme(legend.position = "none")
```


Cumulative sums

```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = Date, y = Cum_Heat),
             color = "red") + 
  geom_path(aes(x = Date, y = Cum_Heat),
            color = "red") + 
    geom_point(aes(x = Date, y = Cum_Cliff),
               color = "blue") + 
  geom_path(aes(x = Date, y = Cum_Cliff),
            color = "blue", alpha = 0.6) +
  labs(title = "Heat and Cliff fall")
```


```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, y = Cum_Heat, color = Season.f)) + 
  geom_path(aes(x = DOS, y = Cum_Heat, color = Season.f)) +
  theme(legend.position = "none")
```



```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, y = Cum_Cliff, color = Season.f)) + 
  geom_path(aes(x = DOS, y = Cum_Cliff, color = Season.f)) +
  theme(legend.position = "none")
```


Cliff fall vs. heat caused deaths

```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = log(Heat + 1), y = log(Cliff + 1)))
```

No obvious correlation between heat and cliff.

In the natural log space

```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = Date, y = log(Cum_Heat + 1)),
             color = "red") + 
  geom_path(aes(x = Date, y = log(Cum_Heat+1)),
            color = "red") + 
    geom_point(aes(x = Date, y = log(Cum_Cliff+1)),
               color = "blue") + 
  geom_path(aes(x = Date, y = log(Cum_Cliff+1)),
            color = "blue", alpha = 0.6) +
  labs(title = "Heat and Cliff fall")
```



```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, 
                 y = log(Cum_Heat+1))) + 
  geom_path(aes(x = DOS,
                y = log(Cum_Heat + 1))) +
  facet_wrap(. ~ Season)

```

Cliff fall

```{r}
ggplot(data = data.1) + 
  geom_point(aes(x = DOS, 
                 y = log(Cum_Cliff+1))) + 
  geom_path(aes(x = DOS,
                y = log(Cum_Cliff + 1))) +
  facet_wrap(. ~ Season)

```

These data are not as clean as the hatchling production data. Maybe I can fit Richard's function to the cumulative sums. 

