---
title: "Turtle deaths modeling"
output: html_notebook
---

I try to model turtle deaths at Raine Island by using a parametric function to determine missing counts.  They are only a few observations each season but the beginning (October 15) and end (April 30) of each season are assumed known. There is a peak during each season. I use modified Richards' function, which was used to model phenology of turtle nesting in Girondot et al. (2007). I extended the modification by Girondot et al. (2007) by creating asymmetrical curves. 

The function in Girondot et al. (2007) has the following form:
S1 <- -S
M1 <- (1 + (2 * exp(K) - 1) * exp((1/S1) * (P - d))) ^ (-1/exp(K))
M2 <- (1 + (2 * exp(K) - 1) * exp((1/S) * (P - d))) ^ (-1/exp(K))
N <- min + (max - min) * (M1 * M2)

d = the number of days from the beginning of nesting season
S > 0 and S1 = -S define the "fatness" of the function
K > 0 defines the "flatness" at the peak of the function
P defines where the peak is relative to the range of d min(d) < P < max(d).
For this analysis, I set P to the middle of the season, that is 
P = floor((max(d) - min(d))/2) because there is not enough data to estimate this parameter. This assumption was bothering me so I'm adding six extra models for allowing P to be estimated as well (2020-06-18). 
min = 0.

Observed counts of deaths are assumed to be normally distributed around this function.

I created six variants of this function by making different assumptions about S and K parameters.
Model 1: S and K are constants (S_K)
Model 2: S is season dependent, whereas K is constant (St_K)
Model 3: S is constant and K for M1 and M2 are different but season independent (S_K1_K2)
Model 4: S is season dependent and K for M1 and M2 are different but season independent (St_K1_K2)
Model 5: absolute values of S and S1 are not the same (creates a non-symmetrical function), whereas K is constant (S1_S2_K)
Model 6: absolute values of S and S1 are not the same and K for M1 and M2 are different but season independent (S1_S2_K1_K2)

For Models 7 through 12, I let P to be another parameter to be estimated. They are named +_P, where + is one of the model names above. 

Varying K by season resulted in non-convergence. Consequently, time dependent K models were not considered. 

Girondot et al. (2007): Modeling approaches to quantify leatherback nesting trends in French Guiana and Suriname. Chelonian Conservation and Biology 6(1): 37-46.

```{r}
rm(list=ls())

library(tidyverse)
library(ggplot2)
library(lubridate)
library(readr)
library(reshape2)
library(jagsUI)
library(bayesplot)
library(loo)

source("Dunstan_functions.R")
```

Get the data. Then, create a dataframe and create Season and Day of Season (DOS) variables.  

```{r}
col.def <- cols(Date = col_date(format = "%m/%d/%Y"),
                Heat = col_integer(),
                Cliff = col_integer(),
                Total = col_integer())

# Begin and end dates for each season
Day.begin <- "10-15"
Day.end <- "04-30"
# define nesting seasons - not the same as calendar years
# also define day of season, starts on 10-15

data.1 <- read_csv(file = "data/Turtle_deaths_with_zeros.csv",
                   col_types = col.def) %>%
  mutate(Year = year(Date),
         DOY = yday(Date),
         Season = ifelse(Date >= paste0(Year, "-", Day.end), 
                         Year, Year-1),
         Season.f = as.factor(Season), 
         DOS = as.numeric(Date - as.Date(paste0(Season, "-", Day.begin)))) %>%  
  group_by(Season.f) %>%
  mutate(Cum_Heat = cumsum(Heat),
         Cum_Cliff = cumsum(Cliff))

# find the number of days between Day1 and Day2
n.days <- as.numeric(as.Date(paste0("2018-", Day.end)) - as.Date(paste0("2017-", Day.begin)))

seasons <- unique(data.1$Season)

data.0 <- data.frame(Season = rep(seasons, each = n.days),
                     DOS = rep(seq(from = 1, to = n.days, by = 1), 
                                    times = length(seasons))) %>% 
  mutate(Date = as.Date(paste0(Season, "-", Day.begin)) + days(DOS))

data.0 %>% left_join(data.1, by = "Date") -> data.0.1

obs.y <- array(NA, c(length(seasons), 2, n.days))

for (i in 1:length(seasons)){
  obs.y[i, ,] <- t(filter(data.0.1, Season.x == seasons[i]) %>% 
                     select(c("Heat", "Cliff")))
  
}

n.seasons <- length(unique(data.1$Season))
max.days <- data.1 %>% 
  group_by(Season.f) %>% 
  summarise(total.days = max(DOS),
            n.days = n()) 

obs.y.1 <- array(data = NA, dim = c(n.seasons, max(max.days$n.days), 2))
obs.t <- matrix(NA, nrow = n.seasons, ncol = max(max.days$n.days))
n.vec <- P <- vector(mode = "numeric", length = n.seasons)
#season.sector.idx <- matrix(NA, nrow = length(Sectors) * length(seasons), ncol = 2)
#max.vec <- vector(mode = "numeric", length = length(seasons))
max.mat <- matrix(NA, nrow = length(seasons), ncol = 2)

season_idx <- cause_idx <- 1
for (season_idx in 1:n.seasons){
  data.1 %>% filter(Season == seasons[season_idx]) -> tmp
  n.vec[season_idx] <- nrow(tmp)
  P[season_idx] <- floor(max(tmp$DOS)/2)
  
  for (cause_idx in 1:2){
    
    if (nrow(tmp) > 0){
      obs.t[season_idx, 1:nrow(tmp)] <- tmp$DOS
      tmp2 <- ifelse(cause_idx == 1,
                     tmp[, "Heat"],
                     tmp[, "Cliff"])
      
      obs.y.1[season_idx, 1:nrow(tmp), cause_idx] <- unlist(tmp2)
      
      max.mat[season_idx, cause_idx] <- ifelse(cause_idx == 1,
                                               max(tmp$Heat, na.rm = T),
                                               max(tmp$Cliff, na.rm = T))
    }
    
  }
}


```

Because no obvious correlation between heat-related and cliff fall deaths, I fit the functions to each cause of death separately. 

Set MCMC parameters

```{r}
MCMC.params <- list(n.chains = 4,
                    n.samples = 50000,
                    n.burnin = 10000,
                    n.thin = 5)

n.per.chain <- (MCMC.params$n.samples - MCMC.params$n.burnin)/MCMC.params$n.thin

model.names <- c("S_K", "St_K", "S_K1_K2",
                 "St_K1_K2", "S1_S2_K", "S1_S2_K1_K2",
                 "S_K_P", "St_K_P", "S_K1_K2_P",
                 "St_K1_K2_P", "S1_S2_K_P", "S1_S2_K1_K2_P")

jags.parameters <- list(c("sigma.y", "S", "K", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S", "K", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S", "K1", "K2", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S", "K1", "K2", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S1", "S2", "K", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S1", "S2", "K1", "K2", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S", "K", "P", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S", "K", "P","max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S", "K1", "K2", "P", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S", "K1", "K2", "P", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S1", "S2", "K", "P", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"),
                        c("sigma.y", "S1", "S2", "K1", "K2", "P", "max",
                          "y", "X", "sigma.max", "deviance", "loglik"))

model.setting.list <- list(jags.parameters = jags.parameters,
                          MCMC.params = MCMC.params,
                          model.names = model.names)

#saveRDS(model.setting.list, file = "RData/model_setting_2020-06-17.rds")

```


# Heat related deaths

```{r}

jags.data.Heat <- list(y = (obs.y.1[,,1]),
                       t = obs.t,
                       P = P,
                       n.vec = n.vec,
                       n.seasons = n.seasons,
                       max.vec = max.mat[,1])

jags.data.Heat_P <- list(y = (obs.y.1[,,1]),
                         t = obs.t,
                         n.vec = n.vec,
                         n.seasons = n.seasons,
                         max.vec = max.mat[,1])

# Convert the data into a vector also and match with the log-likelihood vector
# for computing looic.
data.heat.vector <- as.vector(jags.data.Heat$y) %>% 
  rep(each = MCMC.params$n.chains * n.per.chain)

loo.out.heat <- vector(mode = "list", length = length(model.names))
DIC.heat <- vector(mode = "numeric", length = length(model.names))
max.Rhat.heat <- vector(mode = "numeric", length = length(model.names))

for (k in 1:length(model.names)){
  if (!file.exists(paste0("RData/Richards_Heat_", model.names[[k]], "_norm.rds"))){
    
    if (grep("_P", model.names[[k]]) == 1){
      jags.data <- jags.data.Heat_P
    } else {
      jags.data <- jags.data.Heat
    }
    
    M1.Heat <- jags(jags.data,
                    inits = NULL,
                    parameters.to.save= jags.parameters[[k]],
                    model.file = paste0("models/model_Richards_", model.names[[k]], "_norm.txt"),
                    n.chains = MCMC.params$n.chains,
                    n.burnin = MCMC.params$n.burnin,
                    n.thin = MCMC.params$n.thin,
                    n.iter = MCMC.params$n.samples,
                    DIC = T, parallel=T)
    
    saveRDS(M1.Heat, paste0("RData/Richards_Heat_", model.names[[k]], "_norm.rds"))
    
  } else {
    M1.Heat <- readRDS(paste0("RData/Richards_Heat_", model.names[[k]], "_norm.rds"))
  }

  DIC.heat[k] <- M1.Heat$DIC
  max.Rhat.heat[k] <- data.frame(M1.Heat$summary) %>% select(Rhat) %>% max(na.rm = T)
  
  if (!file.exists(paste0("RData/Richards_Heat_", model.names[[k]], "_norm_LOOIC.rds"))){
    loo.out.heat[[k]] <- compute.looic(loglik = M1.Heat$sims.list$loglik,
                                       data.vector = data.heat.vector,
                                       MCMC.params = MCMC.params)
    
    saveRDS(loo.out.heat[[k]], paste0("RData/Richards_Heat_", model.names[[k]], 
                                      "_norm_LOOIC.rds"))
  } else {
    loo.out.heat[[k]] <- readRDS(paste0("RData/Richards_Heat_", 
                                        model.names[[k]], "_norm_LOOIC.rds"))
  }

}

LOOIC.DIC.Heat <- data.frame(model = c("M1", "M2", "M3", "M4", "M5", "M6",
                                   "M7", "M8", "M9", "M10", "M11", "M12"),
                         LOOIC = unlist(lapply(loo.out.heat, 
                                               FUN = function(x) x$estimates["looic",
                                                                             "Estimate"])),
                         SE = unlist(lapply(loo.out.heat, 
                                            FUN = function(x) x$estimates["looic", "SE"])),
                         DIC = DIC.heat,
                         max.Rhat = max.Rhat.heat) %>%
  arrange(by = LOOIC)

# A quick look at DICs
# DIC.Heat.df <- data.frame(Model = c("M1", "M2", "M3", "M4", "M5", "M6",
#                                     "M7", "M8", "M9", "M10", "M11", "M12"),
#                           DIC = DIC.heat,
#                           max.Rhat = max.Rhat.heat) %>%
#  arrange(by = DIC)
```

Model 10 (St_K1_K2_P) seems to be the best according to LOOIC and DIC values. Convergence was reached for all parameters as well (max.Rhat < 1.1). Pareto k statistics look okay also...

```{r}

best.model.heat.looic <- readRDS(paste0("RData/Richards_Heat_", 
                                        model.names[[10]], "_norm_LOOIC.rds"))

pareto_k_table(best.model.heat.looic)
    
```
There are some bad data points... but we'll use this model for computing the total deaths due to the heat related reasons.

First, take a look at the convergence of the P parameter.

```{r}
best.model.heat <- readRDS(paste0("RData/Richards_Heat_", model.names[[10]], "_norm.rds"))

# According to Rhat statistics, convergence reached for all. 

mcmc_trace(best.model.heat$samples, "P")
```

Look at the density plot:

```{r}
mcmc_dens(best.model.heat$samples, "P")

```

I wonder if we need more MCMC steps... or it's possible that they are different over seasons...


```{r}

X_Heat_St_K1_K2_P.df <- compute.summary.stats_St_K1_K2_P(best.model.heat$samples, 
                                                         seasons, 
                                                         max.days$total.days)

ggplot() + 
  geom_path(data = X_Heat_St_K1_K2_P.df,
            aes(x = DOS, y = Med)) +
  geom_ribbon(data = X_Heat_St_K1_K2_P.df,
              aes(x = DOS, ymin = Low, ymax = High),
              alpha = 0.4) + 
  geom_point(data = data.1,
             aes(x = DOS, y = Heat)) + 
  facet_wrap(.~ as.factor(Season))


```

Now we add all values of Xs for each year

```{r}
X_Heat_St_K1_K2_P.df %>% 
  group_by(Season) %>%
  summarise(total.low = sum(ceiling(Low), na.rm = T),
            total.med = sum(ceiling(Med), na.rm = T),
            total.high = sum(ceiling(High), na.rm = T)) -> total.deaths.heat
```

# Cliff fall
Same can be done for cliff fall deaths:

```{r}
# first just run one at a time
jags.data.Cliff <- list(y = (obs.y.1[,,2]),
                       t = obs.t,
                       P = P,
                       n.vec = n.vec,
                       n.seasons = n.seasons,
                       max.vec = max.mat[,1])

jags.data.Cliff_P <- list(y = (obs.y.1[,,2]),
                          t = obs.t,
                          n.vec = n.vec,
                          n.seasons = n.seasons,
                          max.vec = max.mat[,1])

# Convert the data into a vector also and match with the log-likelihood vector
# for computing looic.
data.cliff.vector <- as.vector(jags.data.Cliff$y) %>% 
  rep(each = MCMC.params$n.chains * n.per.chain)

loo.out.cliff <- vector(mode = "list", length = length(model.names))
DIC.cliff <- vector(mode = "numeric", length = length(model.names))
max.Rhat.cliff <- vector(mode = "numeric", length = length(model.names))

for (k in 1:length(model.names)){
  
  if (!file.exists(paste0("RData/Richards_Cliff_", model.names[[k]], "_norm.rds"))){
    if (grep("_P", model.names[[k]]) == 1){
      jags.data <- jags.data.Cliff_P
    } else {
      jags.data <- jags.data.Cliff
    }
    
    M1.Cliff <- jags(jags.data,
                     inits = NULL,
                     parameters.to.save= jags.parameters[[k]],
                     model.file = paste0("models/model_Richards_", 
                                         model.names[[k]], "_norm.txt"),
                     n.chains = MCMC.params$n.chains,
                     n.burnin = MCMC.params$n.burnin,
                     n.thin = MCMC.params$n.thin,
                     n.iter = MCMC.params$n.samples,
                     DIC = T, parallel=T)
    
    saveRDS(M1.Cliff, paste0("RData/Richards_Cliff_", model.names[[k]], "_norm.rds"))
    
  } else {
    M1.Cliff <- readRDS(paste0("RData/Richards_Cliff_", model.names[[k]], "_norm.rds"))
  }
  
  DIC.cliff[k] <- M1.Cliff$DIC
  max.Rhat.cliff[k] <- data.frame(M1.Cliff$summary) %>% select(Rhat) %>% max(na.rm = TRUE)
  
  if (!file.exists(paste0("RData/Richards_Cliff_", model.names[[k]], "_norm_LOOIC.rds"))){
    #data.cliff.
    
    loo.out.cliff[[k]] <- compute.looic(loglik = M1.Cliff$sims.list$loglik,
                                        data.vector = data.cliff.vector,
                                        MCMC.params = MCMC.params)
    
    saveRDS(loo.out.cliff[[k]], paste0("RData/Richards_Cliff_", model.names[[k]], 
                                       "_norm_LOOIC.rds"))
  } else {
    loo.out.cliff[[k]] <- readRDS(paste0("RData/Richards_Cliff_", 
                                         model.names[[k]], "_norm_LOOIC.rds"))
  }

}

LOOIC.Cliff <- data.frame(model = c("M1", "M2", "M3", "M4", "M5", "M6",
                                    "M7", "M8", "M9", "M10", "M11", "M12"),
                          LOOIC = unlist(lapply(loo.out.cliff, 
                                                FUN = function(x) x$estimates["looic",
                                                                              "Estimate"])),
                          SE = unlist(lapply(loo.out.cliff, 
                                             FUN = function(x) x$estimates["looic", "SE"])),
                          DIC = DIC.cliff,
                          Max.Rhat = max.Rhat.cliff) %>%
  arrange(by = LOOIC)

# A quick look at DICs
# DIC.Cliff.df <- data.frame(Model = c("M1", "M2", "M3", "M4", "M5", "M6",
#                                      "M7", "M8", "M9", "M10", "M11", "M12"),
#                            DIC = DIC.cliff,
#                            Max.Rhat = max.Rhat.cliff) %>%
#   arrange(by = DIC)


```


Model 8 (St_K_P) seems to be the best according to LOOIC and DIC values but the convergence is an issue. I can increase the MCMC sample size later. For now, I'll use the second best (Model 10) for this time. Pareto k statistics look okay also...  

Correction: I rerun M8 with longer chains and got them to converge (Rhat < 1.1) so using M8 for the following. 2020-06-19.

```{r}
best.model.cliff.ID <- 8
best.model.cliff.looic <- readRDS(paste0("RData/Richards_Cliff_", 
                                        model.names[[best.model.cliff.ID]],
                                        "_norm_LOOIC.rds"))

pareto_k_table(best.model.cliff.looic)
    
```
Although some are very bad, the majority are okay. So, I'll go with it. 

We'll use Model 8 for computing the total deaths due to the heat related reasons.

```{r}
best.model.cliff <- readRDS(paste0("RData/Richards_Cliff_", 
                                   model.names[[best.model.cliff.ID]],
                                   "_norm.rds"))

X_Cliff <- compute.summary.stats_St_K_P(best.model.cliff$samples, 
                                                          seasons, 
                                                          max.days$total.days)

ggplot() + 
  geom_path(data = X_Cliff,
            aes(x = DOS, y = Med)) +
  geom_ribbon(data = X_Cliff,
              aes(x = DOS, ymin = Low, ymax = High),
              alpha = 0.4) + 
  geom_point(data = data.1,
             aes(x = DOS, y = Cliff)) + 
  facet_wrap(.~ as.factor(Season))


```


Now we add all values of Xs for each year

```{r}
X_Cliff %>% 
  group_by(Season) %>%
  summarise(total.low = sum(ceiling(Low), na.rm = T),
            total.med = sum(ceiling(Med), na.rm = T),
            total.high = sum(ceiling(High), na.rm = T)) -> total.deaths.cliff

#total.deaths.cliff
```


```{r}
mcmc_trace(best.model.cliff$samples, "P")
```

Start here next... 2020-06-16

Things to do:
1. Add likelihood lines to the models and run all models again DONE
2. Using the likelihood values, compute looic DONE
3. Do model comparison using DIC or looic DONE
4. Consider using mvnormal for heat-related and cliff-fall deaths together N/A

Start here 2020-06-17

The bulk of the analysis is done. So, convert this to a report. I decided not to use mvnormal because of the two variables (heat-related and cliff-fall deaths) are seemingly uncorrelated. 




