---
title: "Turtle deaths modeling"
output: html_notebook
---

I try to model turtle deaths at Raine Island by using a parametric function to determine missing counts.  They are only a few observations each season but the beginning (October 15) and end (April 30) of each season are assumed known. There is a peak during each season.  Different sections (sectors) are treated separately but assumed correlated. The state space of hatchling counts is modeled with an AR(1) process, where the slope (c) is modeled with a discrete Fourier series. 

```{r}
rm(list=ls())

library(tidyverse)
library(ggplot2)
library(lubridate)
library(readr)
library(reshape2)
library(jagsUI)
library(bayesplot)
#source("Dunstan_functions.R")
```

Get the data. Then, create a dataframe containing complete season (DOSeason from 0 to 131) for each season. 

```{r}
col.def <- cols(Date = col_date(format = "%m/%d/%Y"),
                Heat = col_integer(),
                Cliff = col_integer(),
                Total = col_integer())

# Begin and end dates for each season
Day.begin <- "10-15"
Day.end <- "04-30"
# define nesting seasons - not the same as calendar years
# also define day of season, starts on 10-15

data.1 <- read_csv(file = "data/Turtle_deaths_with_zeros.csv",
                   col_types = col.def) %>%
  mutate(Year = year(Date),
         Season = ifelse(Date >= paste0(Year, "-", Day.end), 
                         Year, Year-1),)

# find the number of days between Day1 and Day2
n.days <- as.numeric(as.Date(paste0("2018-", Day.end)) - as.Date(paste0("2017-", Day.begin)))

seasons <- unique(data.1$Season)

data.0 <- data.frame(Season = rep(seasons, each = n.days),
                     DOS = rep(seq(from = 1, to = n.days, by = 1), 
                                    times = length(seasons))) %>% 
  mutate(Date = as.Date(paste0(Season, "-", Day.begin)) + days(DOS))

data.0 %>% left_join(data.1, by = "Date") -> data.0.1

obs.y <- array(NA, c(length(seasons), 2, n.days))

for (i in 1:length(seasons)){
  obs.y[i, ,] <- t(filter(data.0.1, Season.x == seasons[i]) %>% 
                     select(c("Heat", "Cliff")))
  
}

```

I have a tentative model that uses discrete a Fourier series with normal observation models to impute missing data.  Send these data to jags to fit the model

```{r}
# initial value is -2 for unobserved X0, which is pretty close to zero in anti-log space.  
jags.data <- list(y = log(obs.y + 1),
                  mu.X0 = c(-2, -2), #array(data = -2, dim = c(dim(obs.y)[1], dim(obs.y)[2])),
                  day = seq(from = 1, to = n.days, by = 1),
                  n.days = length(seq(from = 1, to = n.days, by = 1)),
                  n.years = dim(obs.y)[1],
                  pi = pi)

jags.params <- c("beta.cos", "beta.sin",
                 #"sigma.X", 
                 "sigma.heat.y", "sigma.cliff.y", "rho",
                 "y", "X", "deviance")

MCMC.params <- list(n.chains = 4,
                    n.samples = 50000,
                    n.burnin = 10000,
                    n.thin = 5)


if (!file.exists(paste0("RData/DFS_X0_imputation_mvnorm_", Day.begin, "_", Day.end, ".rds"))){
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_DFS_imputation_X0_mvnorm.txt',
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, parallel=T)
  
  saveRDS(jm, paste0("RData/DFS_X0_imputation_mvnorm_", Day.begin, "_", Day.end, ".rds"))
  
} else {
  jm <- readRDS(paste0("RData/DFS_X0_imputation_mvnorm_", Day.begin, "_", Day.end, ".rds"))
}


```

put together estimated Xs and observed ys.

```{r}

stats_X0 <- extract.statistics(jm)

#data.3.long <- rbind(data.2, ys.q2.5, ys.q50, ys.q97.5, Xs.q2.5, Xs.q50, Xs.q97.5)
#rm(list = c("jm"))
```

Plot estimates:

```{r}


ggplot() + 
  geom_path(data = stats_X0$Xs_2017R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season)),
            size = 1.2) +
  geom_ribbon(data = stats_X0$Xs_2017R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(`2017R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2017R", 
       x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))
```

Probably shouldn't have these other years... 


```{r}

ggplot() + 
  geom_path(data = stats_X0$Xs_2014R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_X0$Xs_2014R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(`2014R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2014R", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}

ggplot() + 
  geom_path(data = stats_X0$Xs_North ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_X0$Xs_North, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(North),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "North", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}
ggplot() + 
  geom_path(data = stats_X0$Xs_South ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_X0$Xs_South, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(South),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "South", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))


```

```{r}
ggplot() + 
  geom_path(data = stats_X0$Xs_West ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_X0$Xs_West, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(West),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "West", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```

Now that we seem to have pretty "believable" models for all these data, we can go ahead and sum the imputed counts with confidence bands.  

```{r}
all.summary <- summary.stats(stats_X0)

ggplot(data = all.summary) + 
  geom_point(aes(x = Season, y = log10(q50))) + 
  geom_errorbar(aes(x = Season, ymin = log10(q2.5), ymax = log10(q97.5))) +
  facet_wrap(Sector ~., nrow = 2 )
```

We can relax the initial data point as deviates:

```{r}

if (!file.exists(paste0("RData/Fourier_imputation_dnorm_", Day1, "_", Day2, ".rds"))){
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_Fourier_imputation_dnorm.txt',
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, parallel=T)
  
  saveRDS(jm, paste0("RData/Fourier_imputation_dnorm_", Day1, "_", Day2, ".rds"))
  
} else {
  jm <- readRDS(paste0("RData/Fourier_imputation_dnorm_", Day1, "_", Day2, ".rds"))
}


```

put together estimated Xs and observed ys.

```{r}

stats_X0 <- extract.statistics(jm)

#data.3.long <- rbind(data.2, ys.q2.5, ys.q50, ys.q97.5, Xs.q2.5, Xs.q50, Xs.q97.5)
#rm(list = c("jm"))
```

Plot estimates:

```{r}


ggplot() + 
  geom_path(data = stats_X0$Xs_2017R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season)),
            size = 1.2) +
  geom_ribbon(data = stats_X0$Xs_2017R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(`2017R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2017R", 
       x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))
```

Probably shouldn't have these other years... 


```{r}

ggplot() + 
  geom_path(data = stats_X0$Xs_2014R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_X0$Xs_2014R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(`2014R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2014R", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}

ggplot() + 
  geom_path(data = stats_X0$Xs_North ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_X0$Xs_North, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(North),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "North", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}
ggplot() + 
  geom_path(data = stats_X0$Xs_South ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_X0$Xs_South, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(South),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "South", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))


```

```{r}
ggplot() + 
  geom_path(data = stats_X0$Xs_West ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_X0$Xs_West, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(West),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "West", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```

Now that we seem to have pretty "believable" models for all these data, we can go ahead and sum the imputed counts with confidence bands.  

```{r}
all.summary <- summary.stats(stats_X0)

ggplot(data = all.summary) + 
  geom_point(aes(x = Season, y = log10(q50))) + 
  geom_errorbar(aes(x = Season, ymin = log10(q2.5), ymax = log10(q97.5))) +
  facet_wrap(Sector ~., nrow = 2 )
```


What if we have sector and season specific beta parameters for the Fourier function?

```{r}

if (!file.exists(paste0("RData/Fourier_beta_yr_sector_imputation_dnorm_", Day1, "_", Day2, ".rds"))){
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_Fourier_beta_yr_sector_imputation_dnorm.txt',
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, parallel=T)
  
  saveRDS(jm, paste0("RData/Fourier_beta_yr_sector_imputation_dnorm_", Day1, "_", Day2, ".rds"))
  
} else {
  jm <- readRDS(paste0("RData/Fourier_beta_yr_sector_imputation_dnorm_", Day1, "_", Day2, ".rds"))
}


```

put together estimated Xs and observed ys.

```{r}

stats_beta_yr_sector <- extract.statistics(jm)

#data.3.long <- rbind(data.2, ys.q2.5, ys.q50, ys.q97.5, Xs.q2.5, Xs.q50, Xs.q97.5)
#rm(list = c("jm"))
```

Plot estimates:

```{r}


ggplot() + 
  geom_path(data = filter(stats_beta_yr_sector$Xs_2017R, Season == 2017),
            aes(x = DOSeason, y = Xs_q50),
            size = 1.2) +
  geom_ribbon(data = filter(stats_beta_yr_sector$Xs_2017R, Season == 2017), 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(`2017R`)),
             size = 2, shape = 16) +
  labs(title = "2017R", 
       x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))
```


```{r}

ggplot() + 
  geom_path(data = stats_beta_yr_sector$Xs_2014R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_beta_yr_sector$Xs_2014R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(`2014R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2014R", x = paste0("Days since ", Day1), 
       y = "log(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}

ggplot() + 
  geom_path(data = stats_beta_yr_sector$Xs_North ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_beta_yr_sector$Xs_North, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(North),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "North", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}
ggplot() + 
  geom_path(data = stats_beta_yr_sector$Xs_South ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_beta_yr_sector$Xs_South, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(South),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "South", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))


```

```{r}
ggplot() + 
  geom_path(data = filter(stats_beta_yr_sector$Xs_West, Season == 2017),
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = filter(stats_beta_yr_sector$Xs_West, Season == 2017), 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = filter(data.2, Season == 2017),
             aes(x = DOSeason, y = log(West),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "West", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```

Now that we seem to have pretty "believable" models for all these data, we can go ahead and sum the imputed counts with confidence bands.  

```{r}
all.summary <- summary.stats(stats_beta_yr_sector)

ggplot(data = all.summary) + 
  geom_point(aes(x = Season, y = log10(q50))) + 
  geom_errorbar(aes(x = Season, ymin = log10(q2.5), ymax = log10(q97.5))) +
  facet_wrap(Sector ~., nrow = 2 )
```


Also try second-order polynomial here:

```{r}
jags.data <- list(y = log(obs.y),
                  day = data.2$DOSeason,
                  n.steps = dim(obs.y)[3],
                  n.sectors = dim(obs.y)[2],
                  n.years = dim(obs.y)[1],
                  n.days = n.days)

jags.params <- c("sigma.y", "c",
                 "y", "X", "deviance")

MCMC.params <- list(n.chains = 3,
                    n.samples = 100000,
                    n.burnin = 80000,
                    n.thin = 5)


if (!file.exists(paste0("RData/poly_log_imputation_dnorm_", Day1, "_", Day2, ".rds"))){
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_poly_imputation_dnorm.txt',
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, parallel=T)
  
  saveRDS(jm, paste0("RData/poly_log_imputation_dnorm_", Day1, "_", Day2, ".rds"))
  
} else {
  jm <- readRDS(paste0("RData/poly_log_imputation_dnorm_", Day1, "_", Day2, ".rds"))
}

```

```{r}

stats_poly_log <- extract.statistics(jm)

```


```{r}

ggplot() + 
  geom_path(data = filter(stats_poly_log$Xs_2017R, Season == 2017) ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season)),
            size = 1.2) +
  geom_ribbon(data = filter(stats_poly_log$Xs_2017R, Season == 2017),
                aes(x = DOSeason,
                    ymin = Xs_q2.5,
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +
  geom_point(data = filter(data.2, Season == 2017),
             aes(x = DOSeason, y = log(`2017R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2017R", 
       x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))
```

Probably shouldn't have these other years... 


```{r}

ggplot() + 
  geom_path(data = stats_poly_log$Xs_2014R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_poly_log$Xs_2014R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(`2014R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2014R", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```




```{r}

ggplot() + 
  geom_path(data = stats_poly_log$Xs_North ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_poly_log$Xs_North, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(North),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "North", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}
ggplot() + 
  geom_path(data = stats_poly_log$Xs_South ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_poly_log$Xs_South, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(South),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "South", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))


```

```{r}
data.no2014.west <- filter(stats_poly_log$Xs_West, Season != 2014) %>%
  mutate(f.Season = as.factor(Season))
ggplot() + 
  geom_path(data = filter(stats_poly_log$Xs_West, Season != 2014),
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = filter(stats_poly_log$Xs_West, Season != 2014),
                aes(x = DOSeason,
                    ymin = Xs_q2.5,
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +
  geom_point(data = filter(data.2, Season != 2014),
             aes(x = DOSeason, y = log(West),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "West", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```




```{r}
all.summary <- summary.stats(stats_poly_log)

# ggplot(data = all.summary) + 
#   geom_point(aes(x = Season, y = (q50))) + 
#   geom_errorbar(aes(x = Season, ymin = (q2.5), ymax = (q97.5))) +
#   facet_wrap(Sector ~., nrow = 2 )
```



Also try second-order polynomial without log here:

```{r}
jags.data <- list(y = (obs.y),
                  day = data.2$DOSeason,
                  n.steps = dim(obs.y)[3],
                  n.sectors = dim(obs.y)[2],
                  n.years = dim(obs.y)[1],
                  n.days = n.days)
# 
# jags.params <- c("sigma.y", "c",
#                  "y", "X", "deviance")
# 
# MCMC.params <- list(n.chains = 3,
#                     n.samples = 100000,
#                     n.burnin = 80000,
#                     n.thin = 5)


if (!file.exists(paste0("RData/poly_imputation_dnorm_", Day1, "_", Day2, ".rds"))){
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_poly_imputation_dnorm.txt',
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, parallel=T)
  
  saveRDS(jm, paste0("RData/poly_imputation_dnorm_", Day1, "_", Day2, ".rds"))
  
} else {
  jm <- readRDS(paste0("RData/poly_imputation_dnorm_", Day1, "_", Day2, ".rds"))
}

```

```{r}

stats_poly <- extract.statistics(jm)

```


```{r}

ggplot() + 
  geom_path(data = filter(stats_poly$Xs_2017R, Season == 2017) ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season)),
            size = 1.2) +
  geom_ribbon(data = filter(stats_poly$Xs_2017R, Season == 2017),
                aes(x = DOSeason,
                    ymin = Xs_q2.5,
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +
  geom_point(data = filter(data.2, Season == 2017),
             aes(x = DOSeason, y = `2017R`,
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2017R", 
       x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))
```

Probably shouldn't have these other years... 


```{r}

ggplot() + 
  geom_path(data = stats_poly$Xs_2014R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_poly$Xs_2014R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = `2014R`,
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2014R", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```




```{r}

ggplot() + 
  geom_path(data = stats_poly$Xs_North ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_poly$Xs_North, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (North),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "North", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}
ggplot() + 
  geom_path(data = stats_poly$Xs_South ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_poly$Xs_South, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (South),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "South", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))


```

```{r}
data.no2014.west <- filter(stats_poly$Xs_West, Season != 2014) %>%
  mutate(f.Season = as.factor(Season))
ggplot() + 
  geom_path(data = filter(stats_poly$Xs_West, Season != 2014),
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = filter(stats_poly$Xs_West, Season != 2014),
                aes(x = DOSeason,
                    ymin = Xs_q2.5,
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +
  geom_point(data = filter(data.2, Season != 2014),
             aes(x = DOSeason, y = (West),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "West", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```

```{r}
all.summary <- summary.stats(stats_poly)

# ggplot(data = all.summary) + 
#   geom_point(aes(x = Season, y = (q50))) + 
#   geom_errorbar(aes(x = Season, ymin = (q2.5), ymax = (q97.5))) +
#   facet_wrap(Sector ~., nrow = 2 )
```


Next... try the function provided in Girondot et al. (2007).
Try first with non-log counts as it was done in their paper. Change the data format a little so that we are not "imputing" but estimating the parameters of the functions from data.

```{r}
data.1 %>% mutate(DOSeason = as.numeric(Date - as.Date(paste0(Season, "-", Day1)))) %>%
  select(-Date) -> data.3
Sectors <- unique(data.3$Sector)

obs.t <- obs.y.1 <- matrix(NA, nrow = length(Sectors) * length(seasons), ncol = 10)
n.vec <- vector(mode = "numeric", length = length(Sectors) * length(seasons))
season.sector.idx <- matrix(NA, nrow = length(Sectors) * length(seasons), ncol = 2)

max.mat <- matrix(NA, nrow = length(seasons), ncol = length(Sectors))

r <- season_idx <- sector_idx <- 1
for (season_idx in 1:length(seasons)){
  for (sector_idx in 1:length(Sectors)){
    data.3 %>% filter(Season == seasons[season_idx]) %>%
      filter(Sector == Sectors[sector_idx]) -> tmp
    if (nrow(tmp) > 0){
      obs.t[r, 1:nrow(tmp)] <- tmp$DOSeason
      obs.y.1[r, 1:nrow(tmp)] <- tmp$Counts_100m
      n.vec[r] <- nrow(tmp)
      max.mat[season_idx, sector_idx] <- max(tmp$Counts_100m, na.rm = T)      
    }
    season.sector.idx[r,] <- c(season_idx, sector_idx)
    r <- r + 1
  }
}

max.mat[max.mat < 0] <- NA

for (k in 1:ncol(max.mat)){
  max.mat[is.na(max.mat[,k]), k] <- mean(max.mat[, k], na.rm = T)
}

jags.data <- list(y = (obs.y.1),
                  t = obs.t,
                  n.vec = n.vec,
                  idx = season.sector.idx,
                  min = rep(1, length(Sectors)),
                  n = length(n.vec),
                  n.sectors = length(Sectors),
                  n.years = length(seasons),
                  max.mat = max.mat)
# 
jags.params <- c("sigma.y", "P", "S", "K", "max",
                 "y", "X", "sigma.max", "deviance")
# 
# MCMC.params <- list(n.chains = 3,
#                     n.samples = 100000,
#                     n.burnin = 80000,
#                     n.thin = 5)


if (!file.exists(paste0("RData/Girondot_par_estimation_dnorm_", Day1, "_", Day2, ".rds"))){
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_Girondot_Par_estimation_dnorm.txt',
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, parallel=T)
  
  saveRDS(jm, paste0("RData/Girondot_Par_estimation_dnorm_", Day1, "_", Day2, ".rds"))
  
} else {
  jm <- readRDS(paste0("RData/Girondot_Par_estimation_dnorm_", Day1, "_", Day2, ".rds"))
}

```

Need to compute derived (Xs) values
```{r}

Girondot_stats <- list(length = 5)
i <- j <- k <- 1
#k <- 50
for (j in 1:length(Sectors)){
  tmp.mat <- matrix(data = NA, nrow = n.days * length(seasons), ncol = 5)
  c <- 1
  for (i in 1:length(seasons)){
    
    for (k in 1:n.days){
      tmp <- Girondot_fcn(k, 
                          extract.samples("S", jm$samples), 
                          extract.samples("K", jm$samples), 
                          extract.samples(paste0("P[", i, "]"), jm$samples), 
                          jags.data$min[j], 
                          extract.samples(paste0("max[", i, ",", j, "]"), jm$samples)) 
      tmp.mat[c, ] <- c(seasons[i], k, quantile(tmp, c(0.025, 0.5, 0.975)))
      c <- c + 1  
    }
  }
  tmp.df <- data.frame(tmp.mat)
  colnames(tmp.df) <- c("Season", "DOSeason", "Xs_q2.5", "Xs_q50", "Xs_q97.5")
  Girondot_stats[[j]] <- tmp.df
}

names(Girondot_stats) <-paste0("Xs_", Sectors)


```


```{r}

ggplot() + 
  geom_path(data = filter(Girondot_stats$Xs_2017R, Season == 2017) ,
            aes(x = DOSeason, y = Xs_q50),
            size = 1.2) +
  geom_ribbon(data = filter(Girondot_stats$Xs_2017R, Season == 2017),
                aes(x = DOSeason,
                    ymin = Xs_q2.5,
                    ymax = Xs_q97.5),
              alpha = 0.3,
              show.legend = F) +
  geom_point(data = filter(data.2, Season == 2017),
             aes(x = DOSeason, y = `2017R`),
             size = 2, shape = 16) +
  labs(title = "2017R", 
       x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))
```

Probably shouldn't have these other years... 


```{r}

ggplot() + 
  geom_path(data = Girondot_stats$Xs_2014R,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = Girondot_stats$Xs_2014R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = `2014R`,
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2014R", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```




```{r}

ggplot() + 
  geom_path(data = Girondot_stats$Xs_North ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = Girondot_stats$Xs_North, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (North),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "North", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}
ggplot() + 
  geom_path(data = Girondot_stats$Xs_South ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = Girondot_stats$Xs_South, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (South),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "South", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))


```

```{r}
data.no2014.west <- filter(Girondot_stats$Xs_West, Season != 2014) %>%
  mutate(f.Season = as.factor(Season))
ggplot() + 
  geom_path(data = filter(Girondot_stats$Xs_West, Season != 2014),
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = filter(Girondot_stats$Xs_West, Season != 2014),
                aes(x = DOSeason,
                    ymin = Xs_q2.5,
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +
  geom_point(data = filter(data.2, Season != 2014),
             aes(x = DOSeason, y = (West),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "West", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```

```{r}
all.summary <- summary.stats.nolog(Girondot_stats)

# ggplot(data = all.summary) + 
#   geom_point(aes(x = Season, y = (q50))) + 
#   geom_errorbar(aes(x = Season, ymin = (q2.5), ymax = (q97.5))) +
#   facet_wrap(Sector ~., nrow = 2 )
```

Do the same in the log scale
```{r}
jags.data <- list(y = log(obs.y.1),
                  t = obs.t,
                  n.vec = n.vec,
                  idx = season.sector.idx,
                  min = log(rep(1, length(Sectors))),
                  n = length(n.vec),
                  n.sectors = length(Sectors),
                  n.years = length(seasons),
                  max.mat = log(max.mat))
# 
jags.params <- c("sigma.y", "P", "S", "K", "max",
                 "y", "X", "sigma.max", "deviance")
# 
# MCMC.params <- list(n.chains = 3,
#                     n.samples = 100000,
#                     n.burnin = 80000,
#                     n.thin = 5)


if (!file.exists(paste0("RData/Girondot_log_par_estimation_dnorm_", Day1, "_", Day2, ".rds"))){
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_Girondot_Par_estimation_dnorm.txt',
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, parallel=T)
  
  saveRDS(jm, paste0("RData/Girondot_log_Par_estimation_dnorm_", Day1, "_", Day2, ".rds"))
  
} else {
  jm <- readRDS(paste0("RData/Girondot_log_Par_estimation_dnorm_", Day1, "_", Day2, ".rds"))
}

```

Need to compute derived (Xs) values
```{r}

Girondot_stats <- list(length = 5)
i <- j <- k <- 1
#k <- 50
for (j in 1:length(Sectors)){
  tmp.mat <- matrix(data = NA, nrow = n.days * length(seasons), ncol = 5)
  c <- 1
  for (i in 1:length(seasons)){
    
    for (k in 1:n.days){
      tmp <- Girondot_fcn(k, 
                          extract.samples("S", jm$samples), 
                          extract.samples("K", jm$samples), 
                          extract.samples(paste0("P[", i, "]"), jm$samples), 
                          jags.data$min[j], 
                          extract.samples(paste0("max[", i, ",", j, "]"), jm$samples)) 
      tmp.mat[c, ] <- c(seasons[i], k, quantile(tmp, c(0.025, 0.5, 0.975)))
      c <- c + 1  
    }
  }
  tmp.df <- data.frame(tmp.mat)
  colnames(tmp.df) <- c("Season", "DOSeason", "Xs_q2.5", "Xs_q50", "Xs_q97.5")
  Girondot_stats[[j]] <- tmp.df
}

names(Girondot_stats) <-paste0("Xs_", Sectors)


```


```{r}

ggplot() + 
  geom_path(data = filter(Girondot_stats$Xs_2017R, Season == 2017) ,
            aes(x = DOSeason, y = Xs_q50),
            size = 1.2) +
  geom_ribbon(data = filter(Girondot_stats$Xs_2017R, Season == 2017),
                aes(x = DOSeason,
                    ymin = Xs_q2.5,
                    ymax = Xs_q97.5),
              alpha = 0.3,
              show.legend = F) +
  geom_point(data = filter(data.2, Season == 2017),
             aes(x = DOSeason, y = log(`2017R`)),
             size = 2, shape = 16) +
  labs(title = "2017R", 
       x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))
```

Probably shouldn't have these other years... 


```{r}

ggplot() + 
  geom_path(data = Girondot_stats$Xs_2014R,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = Girondot_stats$Xs_2014R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(`2014R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2014R", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```




```{r}

ggplot() + 
  geom_path(data = Girondot_stats$Xs_North ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = Girondot_stats$Xs_North, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(North),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "North", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}
ggplot() + 
  geom_path(data = Girondot_stats$Xs_South ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = Girondot_stats$Xs_South, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = log(South),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "South", x = paste0("Days since ", Day1), 
       y = "Median counts and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))


```



Try it with imputation - doesn't work so well... although Rhat shows convergence, trace plots look quite horrible... 
```{r}

jags.data <- list(y = (obs.y),
                  n.vec = n.vec,
                  min = rep(1, length(Sectors)),
                  n.sectors = length(Sectors),
                  n.years = length(seasons),
                  max.mat = max.mat,
                  n.steps = n.days)
# 

# jags.data <- list(y = (obs.y),
#                   day = data.2$DOSeason,
#                   n.steps = dim(obs.y)[3],
#                   n.sectors = dim(obs.y)[2],
#                   n.years = dim(obs.y)[1],
#                   n.days = n.days)

jags.params <- c("sigma.y", "P", "S", "K", "max",
                 "y", "X", "deviance")
# 
# MCMC.params <- list(n.chains = 3,
#                     n.samples = 100000,
#                     n.burnin = 80000,
#                     n.thin = 5)


if (!file.exists(paste0("RData/Girondot_imputation_dnorm_", Day1, "_", Day2, ".rds"))){
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_Girondot_imputation_dnorm.txt',
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, parallel=T)
  
  saveRDS(jm, paste0("RData/Girondot_imputation_dnorm_", Day1, "_", Day2, ".rds"))
  
} else {
  jm <- readRDS(paste0("RData/Girondot_imputation_dnorm_", Day1, "_", Day2, ".rds"))
}
```

```{r}

stats_Girondot <- extract.statistics(jm)

#data.3.long <- rbind(data.2, ys.q2.5, ys.q50, ys.q97.5, Xs.q2.5, Xs.q50, Xs.q97.5)
#rm(list = c("jm"))
```


```{r}


ggplot() + 
  geom_path(data = stats_Girondot$Xs_2017R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season)),
            size = 1.2) +
  geom_ribbon(data = stats_Girondot$Xs_2017R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (`2017R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2017R", 
       x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))
```

Probably shouldn't have these other years... 


```{r}

ggplot() + 
  geom_path(data = stats_Girondot$Xs_2014R ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_Girondot$Xs_2014R, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (`2014R`),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "2014R", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}

ggplot() + 
  geom_path(data = stats_Girondot$Xs_North ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_Girondot$Xs_North, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (North),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "North", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```



```{r}
ggplot() + 
  geom_path(data = stats_Girondot$Xs_South ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_Girondot$Xs_South, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (South),
                 color = as.factor(Season)),
             size = 2, shape = 16) +
  labs(title = "South", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))


```

```{r}
ggplot() + 
  geom_path(data = stats_Girondot$Xs_West ,
            aes(x = DOSeason, y = Xs_q50,
                color = as.factor(Season))) +
  geom_ribbon(data = stats_Girondot$Xs_West, 
                aes(x = DOSeason,
                    ymin = Xs_q2.5, 
                    ymax = Xs_q97.5,
                    fill = as.factor(Season)),
              alpha = 0.3,
              show.legend = F) +  
  geom_point(data = data.2,
             aes(x = DOSeason, y = (West),
                 color = as.factor(Season)),
             size = 2, shape = 16)+
  labs(title = "West", x = paste0("Days since ", Day1), 
       y = "ln(Median counts) and 95% CI",
       color = "Season") +
  theme(axis.text = element_text(size = 12),
        legend.position = c(0.1, 0.8),
        axis.title = element_text(size = 12))

```

Now that we seem to have pretty "believable" models for all these data, we can go ahead and sum the imputed counts with confidence bands.  

```{r}
all.summary <- summary.stats(stats_X0)

ggplot(data = all.summary) + 
  geom_point(aes(x = Season, y = log10(q50))) + 
  geom_errorbar(aes(x = Season, ymin = log10(q2.5), ymax = log10(q97.5))) +
  facet_wrap(Sector ~., nrow = 2 )
```



