---
title: "Turtle deaths modeling"
output: html_notebook
---

I try to model turtle deaths at Raine Island by using a parametric function to determine missing counts.  They are only a few observations each season but the beginning (October 15) and end (April 30) of each season are assumed known. There is a peak during each season. I use modified Richards' function, which was used to model phenology of turtle nesting in Girondot et al. (2007). I extended the modification by Girondot et al. (2007) by creating asymmetrical curves. 

The function in Girondot et al. (2007) has the following form:
S1 <- -S
M1 <- (1 + (2 * exp(K) - 1) * exp((1/S1) * (P - d))) ^ (-1/exp(K))
M2 <- (1 + (2 * exp(K) - 1) * exp((1/S) * (P - d))) ^ (-1/exp(K))
N <- min + (max - min) * (M1 * M2)

d = the number of days from the beginning of nesting season
S > 0 and S1 = -S define the "fatness" of the function
K > 0 defines the "flatness" at the peak of the function
P defines where the peak is relative to the range of d min(d) < P < max(d).
For this analysis, I set P to the middle of the season, that is 
P = floor((max(d) - min(d))/2) because there is not enough data to estimate this parameter. This assumption was bothering me so I'm adding six extra models for allowing P to be estimated as well (2020-06-18). 
min = 0.

Observed counts of deaths are assumed to be normally distributed around this function.

I created six variants of this function by making different assumptions about S and K parameters.
Model 1: S and K are constants (S_K)
Model 2: S is season dependent, whereas K is constant (St_K)
Model 3: S is constant and K for M1 and M2 are different but season independent (S_K1_K2)
Model 4: S is season dependent and K for M1 and M2 are different but season independent (St_K1_K2)
Model 5: absolute values of S and S1 are not the same (creates a non-symmetrical function), whereas K is constant (S1_S2_K)
Model 6: absolute values of S and S1 are not the same and K for M1 and M2 are different but season independent (S1_S2_K1_K2)

For Models 7 through 12, I let P to be another parameter to be estimated. They are named +_P, where + is one of the model names above. 

Varying K by season resulted in non-convergence. Consequently, time dependent K models were not considered. 

Girondot et al. (2007): Modeling approaches to quantify leatherback nesting trends in French Guiana and Suriname. Chelonian Conservation and Biology 6(1): 37-46.

```{r}
rm(list=ls())

library(tidyverse)
library(ggplot2)
library(lubridate)
library(readr)
library(reshape2)
library(jagsUI)
library(bayesplot)
library(loo)

source("Dunstan_functions.R")
```

Get the data. Then, create a dataframe and create Season and Day of Season (DOS) variables.  

```{r}
col.def <- cols(Date = col_date(format = "%m/%d/%Y"),
                Heat = col_integer(),
                Cliff = col_integer(),
                Total = col_integer())

# Begin and end dates for each season
Day.begin <- "10-15"
Day.end <- "04-30"
# define nesting seasons - not the same as calendar years
# also define day of season, starts on 10-15

data.1 <- read_csv(file = "data/Turtle_deaths_with_zeros.csv",
                   col_types = col.def) %>%
  mutate(Year = year(Date),
         DOY = yday(Date),
         Season = ifelse(Date >= paste0(Year, "-", Day.end), 
                         Year, Year-1),
         Season.f = as.factor(Season), 
         DOS = as.numeric(Date - as.Date(paste0(Season, "-", Day.begin)))) %>%  
  group_by(Season.f) %>%
  mutate(Cum_Heat = cumsum(Heat),
         Cum_Cliff = cumsum(Cliff))

# find the number of days between Day1 and Day2
n.days <- as.numeric(as.Date(paste0("2018-", Day.end)) - as.Date(paste0("2017-", Day.begin)))

seasons <- unique(data.1$Season)

data.0 <- data.frame(Season = rep(seasons, each = n.days),
                     DOS = rep(seq(from = 1, to = n.days, by = 1), 
                                    times = length(seasons))) %>% 
  mutate(Date = as.Date(paste0(Season, "-", Day.begin)) + days(DOS))

data.0 %>% left_join(data.1, by = "Date") -> data.0.1

obs.y <- array(NA, c(length(seasons), 2, n.days))

for (i in 1:length(seasons)){
  obs.y[i, ,] <- t(filter(data.0.1, Season.x == seasons[i]) %>% 
                     select(c("Heat", "Cliff")))
  
}

n.seasons <- length(unique(data.1$Season))
max.days <- data.1 %>% 
  group_by(Season.f) %>% 
  summarise(total.days = max(DOS),
            n.days = n()) 

obs.y.1 <- array(data = NA, dim = c(n.seasons, max(max.days$n.days), 2))
obs.t <- matrix(NA, nrow = n.seasons, ncol = max(max.days$n.days))
n.vec <- P <- vector(mode = "numeric", length = n.seasons)
#season.sector.idx <- matrix(NA, nrow = length(Sectors) * length(seasons), ncol = 2)
#max.vec <- vector(mode = "numeric", length = length(seasons))
max.mat <- matrix(NA, nrow = length(seasons), ncol = 2)

season_idx <- cause_idx <- 1
for (season_idx in 1:n.seasons){
  data.1 %>% filter(Season == seasons[season_idx]) -> tmp
  n.vec[season_idx] <- nrow(tmp)
  P[season_idx] <- floor(max(tmp$DOS)/2)
  
  for (cause_idx in 1:2){
    
    if (nrow(tmp) > 0){
      obs.t[season_idx, 1:nrow(tmp)] <- tmp$DOS
      tmp2 <- ifelse(cause_idx == 1,
                     tmp[, "Heat"],
                     tmp[, "Cliff"])
      
      obs.y.1[season_idx, 1:nrow(tmp), cause_idx] <- unlist(tmp2)
      
      max.mat[season_idx, cause_idx] <- ifelse(cause_idx == 1,
                                               max(tmp$Heat, na.rm = T),
                                               max(tmp$Cliff, na.rm = T))
    }
    
  }
}


```

Because no obvious correlation between heat-related and cliff fall deaths, I fit the functions to each cause of death separately. 

Set MCMC parameters

```{r}
MCMC.params <- list(n.chains = 4,
                    n.samples = 100000,
                    n.burnin = 60000,
                    n.thin = 5)

n.per.chain <- (MCMC.params$n.samples - MCMC.params$n.burnin)/MCMC.params$n.thin

model.names <- c("S_K_norm", "S_K_Pois", "S_K_sigmaYt_norm", 
                 "S_K1_K2_norm", "S_K1_K2_Pois", "S_K1_K2_sigmaYt_norm",
                 
                 "St_K_norm", "St_K_Pois", "St_K_sigmaYt_norm", 
                 "St_K1_K2_norm", "St_K1_K2_Pois", "St_K1_K2_sigmaYt_norm", 
                 
                 "S1_S2_K_norm", "S1_S2_K_Pois", "S1_S2_K_sigmaYt_norm",
                 "S1_S2_K1_K2_norm", "S1_S2_K1_K2_Pois", "S1_S2_K1_K2_sigmaYt_norm",
                 
                 "S_K_P_norm", "S_K_P_Pois", "S_K_P_sigmaYt_norm",
                 "S_K1_K2_P_norm", "S_K1_K2_P_Pois", "S_K1_K2_P_sigmaYt_norm",
                 
                 "St_K_P_norm", "St_K_P_Pois", "St_K_P_sigmaYt_norm",
                 "St_K1_K2_P_norm", "St_K1_K2_P_Pois", "St_K1_K2_P_sigmaYt_norm",
                 
                 "S1_S2_K_P_norm", "S1_S2_K_P_Pois", "S1_S2_K_P_sigmaYt_norm",
                 "S1_S2_K1_K2_P_norm", "S1_S2_K1_K2_P_Pois", "S1_S2_K1_K2_P_sigmaYt_norm")

jags.params.S.K <- c("sigma.y", "S", "K", "max", "P",
                     "y", "X", "sigma.max", "deviance", "loglik")
jags.params.S.K1.K2 <- c("sigma.y", "S", "K1", "K2", "max","P",
                          "y", "X", "sigma.max", "deviance", "loglik")
jags.params.S1.S2.K <- c("sigma.y", "S1", "S2", "K", "max","P",
                         "y", "X", "sigma.max", "deviance", "loglik")
jags.params.S1.S2.K1.K2 <- c("sigma.y", "S1", "S2", "K1", "K2", "max","P",
                             "y", "X", "sigma.max", "deviance", "loglik")

# model.setting.list <- list(jags.parameters = jags.parameters,
#                            MCMC.params = MCMC.params,
#                            model.names = model.names)

#saveRDS(model.setting.list, file = paste0("RData/model_setting_", Sys.Date(), ".rds"))

```


# Heat related deaths

```{r}

jags.data.Heat <- list(y = (obs.y.1[,,1]),
                       t = obs.t,
                       P = P,
                       n.vec = n.vec,
                       n.seasons = n.seasons,
                       max.vec = max.mat[,1])

jags.data.Heat_P <- list(y = (obs.y.1[,,1]),
                         t = obs.t,
                         n.vec = n.vec,
                         n.seasons = n.seasons,
                         max.vec = max.mat[,1])

# Convert the data into a vector also and match with the log-likelihood vector
# for computing looic.
data.heat.vector <- as.vector(jags.data.Heat$y) %>% 
  rep(each = MCMC.params$n.chains * n.per.chain)

loo.out.heat <- vector(mode = "list", length = length(model.names))
DIC.heat <- vector(mode = "numeric", length = length(model.names))
max.Rhat.heat <- vector(mode = "numeric", length = length(model.names))

c <- k <- 1
for (k in 1:length(model.names)){
  if (!file.exists(paste0("RData/Richards_Heat_", model.names[[k]], ".rds"))){
    
    if (length(grep("_P_", model.names[[k]])) == 1){
      jags.data <- jags.data.Heat_P
    } else {
      jags.data <- jags.data.Heat
    }
    
    if (length(grep("S_K_", model.names[[k]])) == 1 | length(grep("St_K_", model.names[[k]])) == 1){
      jags.parameters <- jags.params.S.K
    } else if (length(grep("S_K1_K2_", model.names[[k]])) == 1 | length(grep("St_K1_K2_", model.names[[k]])) == 1){
      jags.parameters <- jags.params.S.K1.K2
    } else if (length(grep("S1_S2_K_", model.names[[k]])) == 1){
      jags.parameters <- jags.params.S1.S2.K
    } else if (length(grep("S1_S2_K1_K2_", model.names[[k]])) == 1){
      jags.parameters <- jags.params.S1.S2.K1.K2
    }
    
    
    M1.Heat <- jags(jags.data,
                    inits = NULL,
                    parameters.to.save= jags.parameters,
                    model.file = paste0("models/model_Richards_", model.names[[k]], ".txt"),
                    n.chains = MCMC.params$n.chains,
                    n.burnin = MCMC.params$n.burnin,
                    n.thin = MCMC.params$n.thin,
                    n.iter = MCMC.params$n.samples,
                    DIC = T, parallel=T)
    
    saveRDS(M1.Heat, paste0("RData/Richards_Heat_", model.names[[k]], ".rds"))
    
  } else {
    M1.Heat <- readRDS(paste0("RData/Richards_Heat_", model.names[[k]], ".rds"))
  }

  DIC.heat[k] <- M1.Heat$DIC
  max.Rhat.heat[k] <- data.frame(M1.Heat$summary) %>% select(Rhat) %>% max(na.rm = T)
  
  if (!file.exists(paste0("RData/Richards_Heat_", model.names[[k]], "_LOOIC.rds"))){
    loo.out.heat[[k]] <- compute.looic(loglik = M1.Heat$sims.list$loglik,
                                       data.vector = data.heat.vector,
                                       MCMC.params = MCMC.params)
    
    saveRDS(loo.out.heat[[k]], paste0("RData/Richards_Heat_", model.names[[k]], 
                                      "_LOOIC.rds"))
  } else {
    loo.out.heat[[k]] <- readRDS(paste0("RData/Richards_Heat_", 
                                        model.names[[k]], "_LOOIC.rds"))
  }

}

LOOIC.DIC.Heat <- data.frame(model = c("M1", "M2", "M3", "M4", "M5", "M6",
                                       "M7", "M8", "M9", "M10", "M11", "M12",
                                       "M13", "M14", "M15", "M16", "M17", "M18",
                                       "M19", "M20", "M21", "M22", "M23", "M24",
                                       "M25", "M26", "M27", "M28", "M29", "M30",
                                       "M31", "M32", "M33", "M34", "M35", "M36"),
                             LOOIC = unlist(lapply(loo.out.heat, 
                                                   FUN = function(x) x$estimates["looic",
                                                                                 "Estimate"])),
                             SE = unlist(lapply(loo.out.heat, 
                                                FUN = function(x) x$estimates["looic", "SE"])),
                             DIC = DIC.heat,
                             max.Rhat = max.Rhat.heat) %>%
  arrange(by = LOOIC)


LOOIC.DIC.Heat
```

Model 29 (St_K1_K2_P_Pois) seems to be the best according to LOOIC and DIC values. Convergence was reached for all parameters as well (max.Rhat < 1.1). Pareto k statistics look okay also...

```{r}
LOOIC.DIC.Heat %>% filter(max.Rhat < 1.1)

```

M20 seems to be good... check Pareto-K

```{r}
best.model.heat.ID <- 20

best.model.heat.looic <- readRDS(paste0("RData/Richards_Heat_", 
                                        model.names[[best.model.heat.ID]], "_LOOIC.rds"))

pareto_k_table(best.model.heat.looic)
    
```

These are good. A bad data point... but we'll use this model for computing the total deaths due to the heat related reasons.

```{r}
model.names[[best.model.heat.ID]]
```

<!-- ```{r} -->
<!-- best.model.heat.ID <- 23 -->

<!-- best.model.heat.looic <- readRDS(paste0("RData/Richards_Heat_",  -->
<!--                                         model.names[[best.model.heat.ID]], "_LOOIC.rds")) -->

<!-- pareto_k_table(best.model.heat.looic) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- model.names[[best.model.heat.ID]] -->
<!-- ``` -->

First, take a look at the convergence of S1, S2, P and K parameters.

```{r}
best.model.heat <- readRDS(paste0("RData/Richards_Heat_", 
                                  model.names[[best.model.heat.ID]], ".rds"))

mcmc_trace(best.model.heat$samples, c("S", "P", "K"))
```

S is not great... 

Look at the density plot:

```{r}
mcmc_dens(best.model.heat$samples, c("S", "P", "K"))


```


<!-- ```{r} -->
<!-- mcmc_trace(best.model.heat$samples, c("K1", "K2")) -->

<!-- ``` -->

I wonder if we need more MCMC steps... 


```{r}

X_Heat_S_K_P <- compute.summary.stats_S_K_P(best.model.heat$samples, 
                                                      seasons, 
                                                      max.days$total.days)

ggplot() + 
  geom_path(data = X_Heat_St_K1_K2_P$X.df,
            aes(x = DOS, y = Med)) +
  geom_ribbon(data = X_Heat_St_K1_K2_P$X.df,
              aes(x = DOS, ymin = Low, ymax = High),
              alpha = 0.4) + 
  geom_point(data = data.1,
             aes(x = DOS, y = Heat)) + 
  facet_wrap(.~ as.factor(Season))


```


# Cliff fall
Same can be done for cliff fall deaths:

```{r}
MCMC.params <- list(n.chains = 4,
                    n.samples = 140000,
                    n.burnin = 100000,
                    n.thin = 5)

# first just run one at a time
jags.data.Cliff <- list(y = (obs.y.1[,,2]),
                       t = obs.t,
                       P = P,
                       n.vec = n.vec,
                       n.seasons = n.seasons,
                       max.vec = max.mat[,1])

jags.data.Cliff_P <- list(y = (obs.y.1[,,2]),
                          t = obs.t,
                          n.vec = n.vec,
                          n.seasons = n.seasons,
                          max.vec = max.mat[,1])

# Convert the data into a vector also and match with the log-likelihood vector
# for computing looic.
data.cliff.vector <- as.vector(jags.data.Cliff$y) %>% 
  rep(each = MCMC.params$n.chains * n.per.chain)

loo.out.cliff <- vector(mode = "list", length = length(model.names))
DIC.cliff <- vector(mode = "numeric", length = length(model.names))
max.Rhat.cliff <- vector(mode = "numeric", length = length(model.names))

for (k in 1:length(model.names)){
  
  if (!file.exists(paste0("RData/Richards_Cliff_", model.names[[k]], ".rds"))){
    if (length(grep("_P_", model.names[[k]])) == 1){
      jags.data <- jags.data.Cliff_P
    } else {
      jags.data <- jags.data.Cliff
    }
    
    M1.Cliff <- jags(jags.data,
                     inits = NULL,
                     parameters.to.save= jags.parameters[[k]],
                     model.file = paste0("models/model_Richards_", 
                                         model.names[[k]], ".txt"),
                     n.chains = MCMC.params$n.chains,
                     n.burnin = MCMC.params$n.burnin,
                     n.thin = MCMC.params$n.thin,
                     n.iter = MCMC.params$n.samples,
                     DIC = T, parallel=T)
    
    saveRDS(M1.Cliff, paste0("RData/Richards_Cliff_", model.names[[k]], ".rds"))
    
  } else {
    M1.Cliff <- readRDS(paste0("RData/Richards_Cliff_", model.names[[k]], ".rds"))
  }
  
  DIC.cliff[k] <- M1.Cliff$DIC
  max.Rhat.cliff[k] <- data.frame(M1.Cliff$summary) %>% select(Rhat) %>% max(na.rm = TRUE)
  
  if (!file.exists(paste0("RData/Richards_Cliff_", model.names[[k]], "_LOOIC.rds"))){
    #data.cliff.
    
    loo.out.cliff[[k]] <- compute.looic(loglik = M1.Cliff$sims.list$loglik,
                                        data.vector = data.cliff.vector,
                                        MCMC.params = MCMC.params)
    
    saveRDS(loo.out.cliff[[k]], paste0("RData/Richards_Cliff_", model.names[[k]], 
                                       "_LOOIC.rds"))
  } else {
    loo.out.cliff[[k]] <- readRDS(paste0("RData/Richards_Cliff_", 
                                         model.names[[k]], "_LOOIC.rds"))
  }

}

LOOIC.DIC.Cliff <- data.frame(model = c("M1", "M2", "M3", "M4", "M5", "M6",
                                        "M7", "M8", "M9", "M10", "M11", "M12",
                                        "M13", "M14", "M15", "M16", "M17", "M18",
                                        "M19", "M20", "M21", "M22", "M23", "M24",
                                        "M25", "M26", "M27", "M28", "M29", "M30",
                                        "M31", "M32", "M33", "M34", "M35", "M36"),
                              LOOIC = unlist(lapply(loo.out.cliff, 
                                                    FUN = function(x) x$estimates["looic",
                                                                                  "Estimate"])),
                              SE = unlist(lapply(loo.out.cliff, 
                                                 FUN = function(x) x$estimates["looic", "SE"])),
                              DIC = DIC.cliff,
                              Max.Rhat = max.Rhat.cliff) %>%
  arrange(by = LOOIC)

LOOIC.DIC.Cliff
```

Some have convergence issues, so find those with Max.Rhat < 1.1

```{r}
LOOIC.DIC.Cliff %>% filter(Max.Rhat < 1.1)

```

Look at some of top ones from this list.  

Pareto_k values show that M18 is a better model than M26 (fewer bad and very bad data points.) So, I'll go with M26 for now.  2020-07-03.
```{r}
best.model.cliff.ID <- 26
best.model.cliff.looic <- readRDS(paste0("RData/Richards_Cliff_", 
                                        model.names[[best.model.cliff.ID]],
                                        "_LOOIC.rds"))

pareto_k_table(best.model.cliff.looic)
    
```

```{r}
model.names[[best.model.cliff.ID]]
```


<!-- ```{r} -->
<!-- best.model.cliff.ID <-  -->
<!-- best.model.cliff.looic <- readRDS(paste0("RData/Richards_Cliff_",  -->
<!--                                         model.names[[best.model.cliff.ID]], -->
<!--                                         "_LOOIC.rds")) -->

<!-- pareto_k_table(best.model.cliff.looic) -->

<!-- ``` -->

Although a couple are very bad, the majority are okay. Take a look at some trace plots:

```{r}
best.model.cliff <- readRDS(paste0("RData/Richards_Cliff_", 
                                   model.names[[best.model.cliff.ID]],
                                   ".rds"))

mcmc_trace(best.model.cliff$samples, c("sigma.y[1]", "sigma.y[2]", "sigma.y[3]", "sigma.y[4]"))
```

```{r}
mcmc_trace(best.model.cliff$samples, c("sigma.y[5]", "sigma.y[6]", "sigma.y[7]", "sigma.y[8]"))

```



```{r}
mcmc_trace(best.model.cliff$samples, c("sigma.y[9]", "sigma.y[10]", "sigma.y[11]", "sigma.y[12]"))

```

These look okay.

```{r}
mcmc_trace(best.model.cliff$samples, c("K"))
```

This looks ok... 


<!-- ```{r} -->
<!-- best.model.cliff.ID <- 10 -->
<!-- best.model.cliff.looic <- readRDS(paste0("RData/Richards_Cliff_",  -->
<!--                                         model.names[[best.model.cliff.ID]], -->
<!--                                         "_LOOIC.rds")) -->

<!-- pareto_k_table(best.model.cliff.looic) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- best.model.cliff <- readRDS(paste0("RData/Richards_Cliff_",  -->
<!--                                    model.names[[best.model.cliff.ID]], -->
<!--                                    ".rds")) -->

<!-- model.names[[best.model.cliff.ID]] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- mcmc_trace(best.model.cliff$samples, c("K1", "K2")) -->
<!-- ``` -->

Not great but better than the others.

```{r}
#mcmc_trace(best.model.cliff$samples, "P")
```

This is a lot better. 

So, I'll use Model 26 for computing the total deaths due to the heat related reasons.

```{r}

X_Cliff <- compute.summary.stats_St_K(samples = best.model.cliff$samples, 
                                      seasons = seasons, 
                                      total.days = max.days$total.days,
                                      P = jags.data.Cliff$P)

ggplot() + 
  geom_path(data = X_Cliff$X.df,
            aes(x = DOS, y = Med)) +
  geom_ribbon(data = X_Cliff$X.df,
              aes(x = DOS, ymin = Low, ymax = High),
              alpha = 0.4) + 
  geom_point(data = data.1,
             aes(x = DOS, y = Cliff)) + 
  facet_wrap(.~ as.factor(Season))


```



Start here next... 2020-06-16

Things to do:
1. Add likelihood lines to the models and run all models again DONE
2. Using the likelihood values, compute looic DONE
3. Do model comparison using DIC or looic DONE
4. Consider using mvnormal for heat-related and cliff-fall deaths together N/A

Start here 2020-06-17

The bulk of the analysis is done. So, convert this to a report. I decided not to use mvnormal because of the two variables (heat-related and cliff-fall deaths) are seemingly uncorrelated. 

2020-07-02
Added Poisson models to both heat and cliff datasets. 

2020-07-02 to 2020-07-03
Added time-dependent sigmaY models. 
Move to Report_Total_Deaths_Estimation.Rmd and its modifications (v2, v3) from here. 




